String Constraints with Regex-Counting and String-Length Solved Efficiently

### 2023-10-18 2023-1-19/sections/algorithms.tex
abstract
review 3 partially
all review finished

This section aims to show how to solve RECL constraints by utilizing CEFA. 
At first, we reduce the satisfiability of RECL constraints to a decision problem defined in the following section. Then we propose an algorithm to solve the decision problem.

@@

Let $e$ be a regular expression over $\Sigma$. Then an occurrence of counting operators in $e$, say $(e')^{\{m,n\}}$ (or $(e')^{\{m, \infty\}}$), is said to be \emph{register-representable} if $(e')^{\{m,n\}}$ (or $(e')^{\{m, \infty\}}$) is not in the scope of a Kleene star, another counting operator, complement, or language difference in $e$. We say that $e$ is \emph{register-representable} if all occurrences of counting operators in $e$ are register-representable. For instance, $(a^{\{2, 6\}} \concat b^*) \cap (a + b)^{\{4, \infty\}}$ is register-representable, while $(a^{\{2, 6\}} \concat b^*) \setminus (a + b)^{\{4, \infty\}}$ and $(a^{\{2, 6\}} \concat b^*)^{\{2,\infty\}}$ are not since $a^{\{2,6\}}$ is in the scope of language difference and the counter operator $\{2, \infty\}$ respectively.

@@

For non-register-representable regular expressions, we first transform them into register-representable regular expressions by unfolding all the non-register-representable occurrences of counting operators. After that, we utilize the aforementioned procedure to construct CEFA. For instance, $(a^{\{2, 6\}} \concat b^*)^{\{2,\infty\}}$ is transformed into $(aa(\varepsilon + a + aa + aaa+aaaa) \concat b^*)^{\{2, \infty\}}$. 
The unfoldings of the inner counting operators of non-register-representable regexes incur an exponential blow-up in the worst case. Nevertheless, those regexes occupy only 5\% of the 48,843 regexes that are collected from the practice (see Section~\ref{sec:bench}). Moreover, the unfoldings are partial in the sense that the outmost counting operators are not unfolded. 
It turns out that our approach can solve almost all the RECL constraints involving these 48,843 regexes, except 181 of them (See Table~\ref{tab:results_regcol}). 
%$(a^{\{2, 6\}} \concat b^*) \setminus (a + b)^{\{4, \infty\}}$ is transformed into $(aa(\varepsilon + a + aa + aaa + aaaa) \concat b^*) \setminus (a+b)(a+b)(a+b)(a+b)(a+b)^*$.

@@

Since
$\sum \limits_{\myvec{v} \in \costset(\cB_i)} \eta(\myvec{v}) \myvec{v} = \sum \limits_{\myvec{v} \in \costset(\cB_i) \setminus \{\myvec{0}\}} \eta(\myvec{v}) \myvec{v}$ for each $i \in [n]$ and $\eta \in \parikh(\cU_{\cB_i})$, it turns out that 
the $\myvec{0}$-labeled transitions in $\cU_{\cB_i}$ do not contribute to the final output $\sum \limits_{\myvec{v} \in \costset(\cB_i)} \eta(\myvec{v}) \myvec{v}$. Therefore, we propose the following simplification technique for the NFA $\cU_{\cB_i}$'s.

>>>>

This section aims to show how to solve RECL constraints by utilizing CEFA. 
At first, we reduce the satisfiability of RECL constraints to a decision problem defined in the following section. Then we propose an decision procedure to solve the problem.

@@

Let $e$ be a regular expression over $\Sigma$. Then an occurrence of counting operators in $e$, say $(e')^{\{m,n\}}$ (or $(e')^{\{m, \infty\}}$), is said to be \emph{register-representable} if $(e')^{\{m,n\}}$ (or $(e')^{\{m, \infty\}}$) is not in the scope of a Kleene star, another counting operator, complement, or language difference in $e$. We say that $e$ is \emph{register-representable} if all occurrences of counting operators in $e$ are register-representable. For instance, $a^{\{2, 6\}} \cap a^{\{4, \infty\}}$ is register-representable, while $\overline{a^{\{2, 6\}}}$ and $(a^{\{2, 6\}})^{\{4,\infty\}}$ are not since $a^{\{2,6\}}$ is in the scope of complement and the counter operator $\{2, \infty\}$ respectively.

@@

For non-register-representable regular expressions, we first transform them into register-representable regular expressions by unfolding all the non-register-representable occurrences of counting operators. After that, we utilize the aforementioned procedure to construct CEFA. For instance, $(a^{\{2, 6\}} \concat b^*)^{\{2,\infty\}}$ is transformed into $(aa(\varepsilon + a + aa + aaa+aaaa) \concat b^*)^{\{2, \infty\}}$. 
The unfoldings of the inner counting operators of non-register-representable regexes incur an exponential blowup in the worst case. Nevertheless, those regexes occupy only 5\% of the 48,843 regexes that are collected from the practice (see Section~\ref{sec:bench}). Moreover, the unfoldings are partial in the sense that the outmost counting operators are not unfolded. 
It turns out that our approach can solve almost all the RECL constraints involving these 48,843 regexes, except 181 of them (See Table~\ref{tab:results_regcol}). 
%$(a^{\{2, 6\}} \concat b^*) \setminus (a + b)^{\{4, \infty\}}$ is transformed into $(aa(\varepsilon + a + aa + aaa + aaaa) \concat b^*) \setminus (a+b)(a+b)(a+b)(a+b)(a+b)^*$.

@@

Since
$\sum \limits_{\myvec{v} \in \costset(\cB_i)} \eta(\myvec{v}) \myvec{v} = \sum \limits_{\myvec{v} \in \costset(\cB_i) \setminus \{\myvec{0}\}} \eta(\myvec{v}) \myvec{v}$ for each $i \in [n]$ and $\eta \in \parikh(\cU_{\cB_i})$, it turns out that 
the $\myvec{0}$-labeled transitions in $\cU_{\cB_i}$ do not contribute to the final output $\sum \limits_{\myvec{v} \in \costset(\cB_i)} \eta(\myvec{v}) \myvec{v}$. Therefore, we propose the following size-reduction technique for the NFA $\cU_{\cB_i}$'s.

### 2023-10-18 2023-1-19/sections/cefa.tex
all review finished

The semantics of CEFA is defined as follows. Let $\aut = (R, Q, \Sigma, \delta, I, F, \alpha)$ be a CEFA. 
%A \emph{configuration} of $\aut$ is a pair $(q, \vec{v})$ where $q \in Q$ and $\vec{v}$ is a vector denoting the values of registers.  
%An \emph{initial configuration} of $\aut$ is $(q_0, \vec{0})$ with $q_0 \in I$, where the value of each register is zero. 
A \emph{run} of $\aut$ on a string $w = a_1 \cdots a_n$ is a sequence $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$ such that $q_0 \in I$ and $q_{i-1} \xrightarrow[\myvec{v_i}]{a_i} q_i$ for each $i \in [n]$. A run $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$ is \emph{accepting} if $q_n \in F$ and $\alpha(\myvec{v'}/R)$ is true, where $\myvec{v'} = \sum \limits_{j \in [n]} \myvec{v_j}$. The vector $\myvec{v'}= \sum \limits_{j \in [n]} \myvec{v_j}$ is called the \emph{cost} of an accepting run $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$. Note that the values of all registers are initiated to zero and updated to $\sum \limits_{j \in [n]} \myvec{v_j}$ after all the transitions in the run are executed. We use $\myvec{v'} \in \aut(w)$ to denote the fact that there is an accepting run of $\aut$ on $w$ whose cost is $\myvec{v'}$.  
We define the semantics of a CEFA $\aut$, denoted by $\Lang(\aut)$, as $\{(w; \myvec{v'}) \mid  \myvec{v'} \in \aut(w)\}$.  In particular, if $I \cap F \neq \emptyset$ and $\alpha(\myvec{0}/R)$ is true, then $(\varepsilon; \myvec{0}) \in \Lang(\aut)$. Moreover, we define the \emph{output} of a CEFA $\aut$, denoted by $\cefaout(\aut)$, as $\{\myvec{v'} \mid  \exists w.\ \myvec{v'} \in \aut(w)\}$.

@@

The \emph{union} of $\aut_1$ and $\aut_2$, denoted by $\aut_1 \cup \aut_2$, is defined in a slightly more involved manner than NFA, as a result of the technical choice that the accepting conditions of CEFA do not distinguish final states. Two fresh auxiliary registers say $r'_1,r'_2 \not \in R_1 \cup R_2$, are introduced so that the accepting condition knows whether a run is from $\aut_1$ (or $\aut_2$). Specifically, $\aut_1 \cup \aut_2 = (R', Q', \Sigma, \delta', I', F', \alpha')$ where 
\begin{itemize}
\item $R' = R_1 \cup R_2 \cup \{r'_1, r'_2\}$, $Q' = Q_1 \cup Q_2 \cup \{q'_0\}$ with $q'_0 \not \in Q_1 \cup Q_2$, $I' = \{q'_0\}$, 
%
\item $\delta'$ is the union of four transitions sets:
\begin{itemize}
  \item $\{(q'_0, a, q'_1, (\myvec{v_1}, \myvec{0}, 1,0)) \mid \exists q_1 \in I_1.\ (q_1, a, q'_1, \myvec{v_1}) \in \delta_1\}$
  \item $\{(q'_0, a, q'_2, (\myvec{0}, \myvec{v_2}, 0, 1)) \mid \exists q_2 \in I_2.\ (q_2, a, q'_2, \myvec{v_2}) \in \delta_2\}$
  \item $\{(q_1, a, q'_1, (\myvec{v_1}, \myvec{0}, 0, 0)) \mid q_1\not\in I_1.\ (q_1, a, q'_1, \myvec{v_1}) \in \delta_1\}$
  \item $\{(q_2, a, q'_2, (\myvec{0}, \myvec{v_2}, 0, 0)) \mid q_2\not\in I_2.\ (q_2, a, q'_2, \myvec{v_2}) \in \delta_2\}$
\end{itemize}
 where $(\myvec{v_1}, \myvec{0}, 1,0)$ is a vector that updates $R_1$ by $\myvec{v_1}$, updates $R_2$ by $\myvec{0}$, and updates $r'_1,r'_2$ by $1,0$ respectively. Similarly for $(\myvec{0}, \myvec{v_2}, 0, 1)$, and so on.

>>>>

The semantics of CEFA is defined as follows. Let $\aut = (R, Q, \Sigma, \delta, I, F, \alpha)$ be a CEFA. 
%A \emph{configuration} of $\aut$ is a pair $(q, \vec{v})$ where $q \in Q$ and $\vec{v}$ is a vector denoting the values of registers.  
%An \emph{initial configuration} of $\aut$ is $(q_0, \vec{0})$ with $q_0 \in I$, where the value of each register is zero. 
A \emph{run} of $\aut$ on a string $w = a_1 \cdots a_n$ is a sequence $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$ such that $q_0 \in I$ and $q_{i-1} \xrightarrow[\myvec{v_i}]{a_i} q_i$ for each $i \in [n]$. A run $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$ is \emph{accepting} if $q_n \in F$ and $\alpha(\myvec{v'}/R)$ is true, where $\myvec{v'} = \sum \limits_{j \in [n]} \myvec{v_j}$. The vector $\myvec{v'}= \sum \limits_{j \in [n]} \myvec{v_j}$ is called the \emph{cost} of an accepting run $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$. Note that the values of all registers are initiated to zero and updated to $\sum \limits_{j \in [n]} \myvec{v_j}$ after all the transitions in the run are executed. We use $\myvec{v'} \in \aut(w)$ to denote the fact that there is an accepting run of $\aut$ on $w$ whose cost is $\myvec{v'}$.  
We define the semantics of a CEFA $\aut$, denoted by $\Lang(\aut)$, as $\{(w; \myvec{v'}) \mid  \myvec{v'} \in \aut(w)\}$.  In particular, if $I \cap F \neq \emptyset$ and $\alpha[\myvec{0}/R]$ is true, then $(\varepsilon; \myvec{0}) \in \Lang(\aut)$. Moreover, we define the \emph{output} of a CEFA $\aut$, denoted by $\cefaout(\aut)$, as $\{\myvec{v'} \mid  \exists w.\ \myvec{v'} \in \aut(w)\}$.

@@

The \emph{union} of $\aut_1$ and $\aut_2$, denoted by $\aut_1 \cup \aut_2$ \sout{is defined in a slightly more involved manner than NFA, as a result of the technical choice that the accepting conditions of CEFA do not distinguish final states}. Two fresh auxiliary registers say $r'_1,r'_2 \not \in R_1 \cup R_2$, are introduced so that the accepting condition knows whether a run is from $\aut_1$ (or $\aut_2$). Specifically, $\aut_1 \cup \aut_2 = (R', Q', \Sigma, \delta', I', F', \alpha')$ where 
\begin{itemize}
\item $R' = R_1 \cup R_2 \cup \{r'_1, r'_2\}$, $Q' = Q_1 \cup Q_2 \cup \{q'_0\}$ with $q'_0 \not \in Q_1 \cup Q_2$, $I' = \{q'_0\}$, 
%
\item $\delta'$ is the union of four transitions sets:
\begin{itemize}
  \item $\{(q'_0, a, q'_1, (\myvec{v_1}, \myvec{0}, 1,0)) \mid \exists q_1 \in I_1.\ (q_1, a, q'_1, \myvec{v_1}) \in \delta_1\}$
  \item $\{(q'_0, a, q'_2, (\myvec{0}, \myvec{v_2}, 0, 1)) \mid \exists q_2 \in I_2.\ (q_2, a, q'_2, \myvec{v_2}) \in \delta_2\}$
  \item $\{(q_1, a, q'_1, (\myvec{v_1}, \myvec{0}, 0, 0)) \mid q_1\not\in I_1.\ (q_1, a, q'_1, \myvec{v_1}) \in \delta_1\}$
  \item $\{(q_2, a, q'_2, (\myvec{0}, \myvec{v_2}, 0, 0)) \mid q_2\not\in I_2.\ (q_2, a, q'_2, \myvec{v_2}) \in \delta_2\}$
\end{itemize}
 where $(\myvec{v_1}, \myvec{0}, 1,0)$ is a vector that updates $R_1$ by $\myvec{v_1}$, updates $R_2$ by $\myvec{0}$, and updates $r'_1,r'_2$ by $1,0$ respectively. Similarly for $(\myvec{0}, \myvec{v_2}, 0, 1)$, and so on.

### 2023-10-18 2023-1-19/sections/implementation.tex
all review finished

%!TEX root = ../main.tex
%\documentclass{standalone}
%\begin{document}
In this section, we list our experimental results. We first present the overall experiment results on RegCoL and AutomatArk benchmark suites. Then we present experiment results of the large-counting part and evaluate the simplification techniques. \denghang{The goal of our experiments is to answer the following questions:
\begin{itemize}
  \item [Q1:] Are our decision procedure more efficient than the state-of-the-art string constraint solvers on RECL constraints?
  \item [Q2:] Why our decision procedure is more efficient?
  \item [Q3:] Are the size-reduction techniques helpful to improve the performance?
\end{itemize}
}

@@

Our experiments utilize two benchmark suites, namely, \emph{RegCoL} and \emph{AutomatArk}. There are 48,843 instances in total, and all benchmark instances are in the SMTLIB2 format.
Moreover, it turns out that only 5\% of regexes among the 48,843 instances are non-register-representable (see Section~\ref{subsec:regex2cefa}).

@@

\medskip
\noindent
\emph{AutomatArk benchmark suite.}
This benchmark suite is adapted from the AutomatArk suite \cite{z3str3re} by picking out the string constraints containing counting operators. We also add the length constraint $|x| > 10$ for each string variable $x$. There are 8,215 instances in the AutomatArk suite.
Note that the original AutomatArk benchmark suite \cite{z3str3re} includes 19,979 instances, which are conjunctions of regular membership queries generated out of regular expressions in \cite{automatark}.
%Only 8,751 contain counting operators out of 19,979 instances.

%%%%%%%%%%%%%%%%%% the benchmark suite written by Denghang %%%%%%%%%%%
%%%%%%%%%%%%%%%%%% the benchmark suite written by Denghang %%%%%%%%%%%

@@

\vspace{-2mm}
\subsection{Experiment Results}
\vspace{-1mm}

@@

\medskip
\noindent
\emph{Overall evaluation.} 
The experiment results can be found in Table~\ref{tab:results_regcol}. Note that we take the result of CVC5 as the ground truth\footnote{Initially,  we used the majority vote of the results of the solvers as the ground truth. Nevertheless, on some problem instances, all the results of the three solvers in the Z3 family are wrong (after manual inspection), thus failing this approach on these instances.}, \denghang{and the results distinct to the ground truth are called \emph{soundness errors} in the result table.} We can see that $\ostrichrecl$ solves almost all 48,843 instances, except 182 of them, that is, it solves \textbf{48,662} instances correctly. The number is %3,994/2,272/24,577/12,832/1,091/7,092 
\textbf{3,908/1,111/12,838/2,306/2,396 more} than the number of instances solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively.
%      
%A soundness error is reported if the result of a solver for an instance is different from CVC5. 
% their results on some instances are inconsistent with the results of the other three solvers, i.e. CVC5, OSTRICH, and OSTRICH$^{\rm RECL}$.
%
Moreover, $\ostrichrecl$ is the second fastest solver, whose average time on each instance is close to the fastest solver Z3str3RE (\textbf{1.93s}/1.62s).

%
\begin{table}
  \vspace{-4mm}
  \centering
  \begin{subtable}{0.513\textwidth}
      \centering
      \import{tables}{table_regcol.tex}
      \caption{Overall evaluation, with timeout = 60 seconds.}
      \label{tab:results_regcol}
  \end{subtable}
  \begin{subtable}{0.477\textwidth}
      \centering
      \import{tables}{table_simp.tex}
      \caption{Evaluation of the automata size-reduction techniques, with timeout = 60 seconds.}
      \label{tab:results_simp}
  \end{subtable}
  \vspace{-2mm}
  \caption{Experiment results}
  \vspace{-8mm}
\end{table}
% \begin{table}[ht]
%   \vspace{-5mm}
%   \begin{center}
%     \import{tables}{table_regcol.tex}
%   \end{center}
%   \caption{Overall Experiment results, with timeout = 60 seconds}
%   \label{tab:results_regcol}
%   \vspace{-9mm}
% \end{table}

\smallskip
\noindent
\emph{Evaluation on problem instances with large bounds.}
%
We extract the 1,969 problem instances with large counting bounds (greater than or equal to $50$) from the RegCoL and AutomatArk benchmark suites.  
%regex with upper bound greater than 50 and generate a \emph{`large-counting'} benchmark suite (1,969 instances in total). 
Moreover, in order to test the performance of the solvers on string constraints with large length bounds as well, we increase the length bound to $200$, that is, $|x| > 200$.
%
\begin{figure}[ht]
  \centering
  \begin{subfigure}[t]{0.49\textwidth}
    \centering\vskip 0pt
    \includegraphics[width=1\textwidth]{counting_distribution.png}  
    \caption{Distribution of problem instances w.r.t. counting bounds}  
    \label{fig:count_distri}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \centering\vskip 0pt
    \import{tables}{table_large_count.tex}
    \vspace{4.5mm}
    \caption{Experiment results for large bounds, \\with timeout = 60 seconds.}
    \label{fig:table_large_count}
  \end{subfigure}
  \vspace{-2mm}
  \caption{Distribution of counting bounds and experiment results for large bounds}
\vspace{-5mm}
\end{figure}

@@

We then evaluate the performance of $\ostrichrecl$ on these 1,969 instances. 
The experiment results are put in Fig~\ref{fig:table_large_count}. From the results, $\ostrichrecl$ solves \textbf{1,873} instances correctly, which is \textbf{947/278/563/637/523 more} than that solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively. Moreover, $\ostrichrecl$ is \textbf{ 6.79/2.88/2.61/5.27/3.95} times faster than CVC5/ Z3str3RE/Z3str3/Z3seq/OSTRICH respectively.

\medskip
\noindent 
\emph{Evaluation of the effectiveness of the automata size-reduction techniques.}
We also do experiments to evaluate the effectiveness of the automata size-reduction techniques, that is, Step 2 in Section~\ref{subsec:cefadec}. 
Let us use $\ostrichrecl_{\rm -SIMP}$ to denote $\ostrichrecl$ with the size-reduction techniques removed. 
Meanwhile, we compare $\ostrichrecl$ and $\ostrichrecl_{\rm -SIMP}$ against $\ostrichrecl_{\rm NUXMV}$, where the $\cefadec$ problem is solved by the \textsc{nuXmv} model checker \cite{nuxmv}, instead of the procedure in Section~\ref{subsec:cefadec}. 
The experiment results are put in Table~\ref{tab:results_simp}. 
%$\ostrichrecl_{\rm -SIMP}$ is the tool without simplification techniques. $\ostrichrecl_{\rm NUXMV}$ is the tool using \textsc{nuXmv}-based techniques. 
From the results, $\ostrichrecl$ solves \textbf{1,503 more} instances and is \textbf{2.21} times faster than $\ostrichrecl_{\rm -SIMP}$, which shows that the size-reduction techniques indeed play a vital role in the performance improvement. Moreover, $\ostrichrecl$ solves \textbf{1,798 more} instances and is \textbf{3.13} times faster than $\ostrichrecl_{\rm NUXMV}$.

>>>>

%!TEX root = ../main.tex
%\documentclass{standalone}
%\begin{document}
In this section, we list our experimental results. We first present the overall experiment results on RegCoL and AutomatArk benchmark suites. Then we present experiment results of the large-counting part and evaluate the simplification techniques. \denghang{The goal of our experiments is to answer the following questions:
\begin{itemize}
  \item [Q1:] Are our decision procedure more efficient than the state-of-the-art string constraint solvers on RECL constraints? (Sec \ref{subsec:overall_eval})
  \item [Q2:] Why our decision procedure is more efficient?(Sec \ref{subsec:large_bounds_eval})
  \item [Q3:] Are the size-reduction techniques practical?(Sec \ref{subsec:size_reduction_eval})
  \item [Q4:] How effective is the algorithm for solving the nonemptiness problem compared to the case that uses the nuXmv model checker? (Sec \ref{subsec:size_reduction_eval})
\end{itemize}
}

@@

Our experiments utilize two benchmark suites, namely, \emph{RegCoL} and \emph{AutomatArk}. \denghang{Other industrial benchmarks suites are not utilized because they contains no counting operator.} There are 48,843 instances in total, and all benchmark instances are in the SMTLIB2 format.
Moreover, it turns out that only 5\% of regexes among the 48,843 instances are non-register-representable (see Section~\ref{subsec:regex2cefa}).

@@

\medskip
\noindent
\emph{AutomatArk benchmark suite.}
This benchmark suite is adapted from the AutomatArk suite \cite{z3str3re} by picking out the string constraints containing counting operators. We also add the length constraint $|x| > 10$ for each string variable $x$. There are 8,215 instances in the AutomatArk suite.
Note that the original AutomatArk benchmark suite \cite{z3str3re} includes 19,979 instances, which are conjunctions of regular membership queries generated out of regular expressions in \cite{automatark}.

\medskip
\noindent
\denghang{
\emph{Large-Bounds part.} This benchmark suite is composed of 1,969 problem instances with large counting bounds (greater than or equal to $50$) from RegCoL and AutomatArk benchmark suites.  
%regex with upper bound greater than 50 and generate a \emph{`large-counting'} benchmark suite (1,969 instances in total). 
Moreover, in order to test the performance of the solvers on string constraints with large length bounds as well, we increase the length bound to $200$, that is, $|x| > 200$.
}
%%%%%%%%%%%%%%%%%% the benchmark suite written by Denghang %%%%%%%%%%%
%%%%%%%%%%%%%%%%%% the benchmark suite written by Denghang %%%%%%%%%%%

@@

\vspace{-2mm}
\subsection{Overall Evaluation}\label{subsec:overall_eval}
\vspace{-1mm}

@@

% \medskip
% \noindent
% \emph{Overall evaluation.} 
The experiment results can be found in Table~\ref{tab:results_regcol}. Note that we take the result of CVC5 as the ground truth\footnote{Initially,  we used the majority vote of the results of the solvers as the ground truth. Nevertheless, on some problem instances, all the results of the three solvers in the Z3 family are wrong (after manual inspection), thus failing this approach on these instances.}, \denghang{and the results distinct to the ground truth are called \emph{soundness errors} in the result table.} We can see that $\ostrichrecl$ solves almost all 48,843 instances, except 182 of them, that is, it solves \textbf{48,662} instances correctly. The number is %3,994/2,272/24,577/12,832/1,091/7,092 
\textbf{3,908/1,111/12,838/2,306/2,396 more} than the number of instances solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively.
%      
%A soundness error is reported if the result of a solver for an instance is different from CVC5. 
% their results on some instances are inconsistent with the results of the other three solvers, i.e. CVC5, OSTRICH, and OSTRICH$^{\rm RECL}$.
%
Moreover, $\ostrichrecl$ is the second fastest solver, whose average time on each instance is close to the fastest solver Z3str3RE (\textbf{1.93s}/1.62s). \denghang{To answer \textbf{Q1}, \textbf{$\ostrichrecl$ is the second fastest solver on these benchmark suites, and it solves the most number of instances}.}

%
\begin{table}
  \vspace{-4mm}
  \centering
  \begin{subtable}{0.513\textwidth}
      \centering
      \import{tables}{table_regcol.tex}
      \caption{Overall evaluation, with timeout = 60 seconds.}
      \label{tab:results_regcol}
  \end{subtable}
  \begin{subtable}{0.477\textwidth}
      \centering
      \import{tables}{table_simp.tex}
      \caption{Evaluation of the size-reduction techniques, with timeout = 60 seconds.}
      \label{tab:results_simp}
  \end{subtable}
  \vspace{-2mm}
  \caption{Overall evaluation and the evaluation of the size-reduction techniques}
  \vspace{-8mm}
\end{table}
% \begin{table}[ht]
%   \vspace{-5mm}
%   \begin{center}
%     \import{tables}{table_regcol.tex}
%   \end{center}
%   \caption{Overall Experiment results, with timeout = 60 seconds}
%   \label{tab:results_regcol}
%   \vspace{-9mm}
% \end{table}

\subsection{Evaluation on problem instances with large bounds}\label{subsec:large_bounds_eval}
% \smallskip
% \noindent
% \emph{Evaluation on problem instances with large bounds.}
%
\sout{We extract the 1,969 problem instances with large counting bounds (greater than or equal to $50$) from the RegCoL and AutomatArk benchmark suites.  
%regex with upper bound greater than 50 and generate a \emph{`large-counting'} benchmark suite (1,969 instances in total). 
Moreover, in order to test the performance of the solvers on string constraints with large length bounds as well, we increase the length bound to $200$, that is, $|x| > 200$.}
%
\begin{figure}[ht]
  \centering
  \begin{subfigure}[t]{0.49\textwidth}
    \centering\vskip 0pt
    \includegraphics[width=1\textwidth]{counting_distribution.png}  
    \caption{Distribution of problem instances w.r.t. counting bounds}  
    \label{fig:count_distri}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \centering\vskip 0pt
    \import{tables}{table_large_count.tex}
    \vspace{4.5mm}
    \caption{Experiment results on Large-Bounds part, with timeout = 60 seconds.}
    \label{fig:table_large_count}
  \end{subfigure}
  \vspace{-2mm}
  \caption{Distribution of counting bounds and experiment results for large bounds}
\vspace{-5mm}
\end{figure}

@@

We evaluate the performance of $\ostrichrecl$ on 1,969 instances of the Large-Bounds part. 
The experiment results are put in Fig~\ref{fig:table_large_count}. From the results, $\ostrichrecl$ solves \textbf{1,873} instances correctly, which is \textbf{947/278/563/637/523 more} than that solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively. Moreover, $\ostrichrecl$ is \textbf{ 6.79/2.88/2.61/5.27/3.95} times faster than CVC5/ Z3str3RE/Z3str3/Z3seq/OSTRICH respectively. \denghang{To answer \textbf{Q2}, we can see that $\ostrichrecl$ is far more efficient than the other solvers on problem instances with large bounds. \textbf{It is because the complexity in our decision procedure is irrelevant to counting bounds}, while the complexity of other solvers is linear or even exponential to counting bounds.}

\subsection{Evaluation of the effectiveness of size-reduction techniques}\label{subsec:size_reduction_eval}
% \medskip
% \noindent 
% \emph{Evaluation of the effectiveness of the automata size-reduction techniques.}
We also do experiments to evaluate the effectiveness of the automata size-reduction techniques, that is, Step 2 in Section~\ref{subsec:cefadec}. 
Let us use $\ostrichrecl_{\rm -SIMP}$ to denote $\ostrichrecl$ with the size-reduction techniques removed. 
Meanwhile, we compare $\ostrichrecl$ and $\ostrichrecl_{\rm -SIMP}$ against $\ostrichrecl_{\rm NUXMV}$, where the $\cefadec$ problem is solved by the \textsc{nuXmv} model checker \cite{nuxmv}, instead of the procedure in Section~\ref{subsec:cefadec}. 
The experiment results are put in Table~\ref{tab:results_simp}. 
%$\ostrichrecl_{\rm -SIMP}$ is the tool without simplification techniques. $\ostrichrecl_{\rm NUXMV}$ is the tool using \textsc{nuXmv}-based techniques. 
From the results, $\ostrichrecl$ solves \textbf{1,503 more} instances and is \textbf{2.21} times faster than $\ostrichrecl_{\rm -SIMP}$. \denghang{To answer \textbf{Q3}, \textbf{the size-reduction techniques indeed play a vital role in the performance improvement}}. Moreover, $\ostrichrecl$ solves \textbf{1,798 more} instances and is \textbf{3.13} times faster than $\ostrichrecl_{\rm NUXMV}$. \denghang{To answer \textbf{Q4}, \textbf{our algorithm for solving the nonemptiness problem is much better than the \textsc{nuXmv}-based techniques}}.
